---
title: "Number and Brightness Image Analysis in R"
author: "Rory Nolan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Number and Brightness analysis in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

### Loading the library

First, let's load the `nandb` library and `magrittr`, which we'll use.
```{r load library}
library("nandb")
library("tidyverse", quietly = TRUE)
```

### Reading and Displaying Images

Now let's read in an image:
```{r read image}
img <- system.file("extdata", "FKBP-mClover_before_0.5nM_AP1510.tif", 
                   package = "nandb") %>% ReadImageData
dim(img)  # get img's dimensions
```
So we can see that `img` is a 3-dimensional array of 50 slices each of which is a 256x256 image. We can view the first slice:
```{r}
EBImage::display(EBImage::normalize(img[, , 1]), method = "raster")
```

### Preprocessing, Brightness Calculation and Post-Filtering

Now we can calculate the brightness image based on this image series, with an exponential filtering detrend with a time constant of (say) 10 frames and using the Huang thresholding method via 
```{r Brightness calculation with bleaching correction}
img_brightness <- Brightness(img, tau = 10, mst = "Huang")
```
and we can see the (median filtered) version of the output. 
```{r Brightness Plot, fig.width=4.5}
img_brightness %>% MedianFilterB %>% 
  MatrixRasterPlot(scale.name = "brightness", limits = c(0.9, 1.5))
```

### Saving Text Images

We can save the brightness image (let's do it without median filtering this time). It saves as a text csv file (not as a tiff image, since the brightness values can be any non-negative real number, but tiff images can only store pixel values as integers).
```{r Brightness write to txt}
WriteImageTxt(img_brightness, "BrightnessImage")
list.files(pattern = "\\.csv$")  # check that the csv file was written to the folder
```
```{r, echo=FALSE, results="hide"}
file.remove("BrightnessImage.csv")  # remove that file (clean up)
```

### Studying the Distribution of Brightnesses

We can take a look at the distribution of brightnesses in that image:
```{r brightness density}
db <- density(img_brightness, na.rm = TRUE) %>% 
  {data.frame(x = .$x, y = .$y)}
ggplot(db, aes(x, y)) + geom_line() + 
  labs(x = "brightness", y = "frequency")
```

and we can compare it to the distribution of brightnesses from immobile particles:
```{r immobile plot}
immobile_brightnesses <- replicate(512^2, rpois(50, 50) %>% {var(.) / mean(.)})
di <- density(immobile_brightnesses) %>% {data.frame(x = .$x, y = .$y)}
rbind(mutate(db, mobility = "mobile"), mutate(di, mobility = "immobile")) %>%     mutate(mobility = factor(mobility)) %>% 
  ggplot(aes(x, y)) + geom_line(aes(colour = mobility)) + 
  labs(x = "brightness", y = "frequency")
  
```

### Calculating the Brightness of a Monomer

Say we know that our monomeric brightness is 1.10. Then we can create a plot of all kmers present:
```{r kmer plot}
KmerPlot(img_brightness, 1.10)
```

Instead, say that you know that you have only monomers (no dimers, trimers, etc) in your images but you don't know what your monomeric brightness is. Then you can use 
```{r}
# this
```
However, if you don't have an image of monomers only, then I can't help you to determine the brightness of your monomer.
